---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

```{r load, include=FALSE}
marginalization = read_csv("../data/marginalization.csv", col_types=cols())
```

The Ontario Marginaliaztion Index (ON-Marg) is a Public Health Ontario project combining numerous demographic indicators into four measures of community marginalization: residential instability, material deprivation, economic dependency, and ethnic concentration^1^. These measures are linked to geography via the 2016 Canada Census via various geographical levels ranging from dissemination areas (the smallest standard geographic areas) to Local Health Integration Networks (LHINs) covering multiple municipalities^2^. We chose to use the census subdivision (CSD) level for our analysis, which correspond to "municipalities (as determined by provincial/territorial legislation) or areas treated as municipal equivalents". We reasoned that for a province-wide vaccination strategy, CSDs strike a good balance between granularity in identifying high-risk groups and controlling the administrative overhead of managing the vaccination campaign.

We related the ON-Marg index to COVID-19 impact using Public Health Ontario's weekly epidemiologic summaries. These summaries contain a comparison of the rate and number of confirmed COVID-19 cases for each quintile of neighbourhood deprivation as defined by the ON-Marg index's material deprivation dimension^3^. These relative COVID-19 rates are expressed in terms of case rate per 100,000 people, so we used census data from the 2016 Canada Census^4^ to convert this to an absolute measure of the mean distribution of cases across each quintile.

```{r risk, echo=FALSE}
marginalization %>%
  select(Deprivation, DeprivationModifier) %>%
  distinct() %>%
  ggplot(aes(x=Deprivation, y=DeprivationModifier, label=round(DeprivationModifier, 3))) +
  geom_bar(stat="identity") +
  geom_text(vjust=-0.5) +
  xlab("Material Deprivation Quintile") +
  ylab("Proportion of Confirmed COVID-19 Cases") +
  ggtitle("Distribution of COVID-19 Cases in Material Deprivation Quintiles")
```

These proportions serve as each material deprivation quintile's weight in the simulation's case exposure selection.

**References:**  
^1^ Matheson FI, van Ingen T. 2016 Ontario marginalization index. Toronto, ON: St. Michael’s Hospital; 2018. Joint publication with Public Health Ontario.  
^2^ Matheson FI, van Ingen T. 2016 Ontario marginalization index: user guide. Toronto, ON: St. Michael’s Hospital; 2018. Joint publication withPublic Health Ontario.  
^3^ Ontario Agency for Health Protection and Promotion (Public Health Ontario). Weekly epidemiologic summary: COVID-19 in Ontario, May 31, 2020 to February 13, 2020. Toronto, ON: Queen’s Printer for Ontario; 2021.  
^4^ "2016 Canada Census.” https://censusmapper.ca. Statistics Canada. Accessed October 10, 2020. https://censusmapper.ca