---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

```{r mortality, include=FALSE}
cases = read_csv("../data/conposcovidloc.csv") %>%
  filter(!is.na(Age_Group)) %>%
  mutate(Age_Group = replace(Age_Group, Age_Group %in% c("80s", "90s"), "80+"))
```

It is well-established in Ontario that COVID-19 mortality rates are variable and highly dependent on age^1^. This suggested that it would be prudent to establish different COVID-19 mortality rates for different age groups.  

We used the Government of Ontario's daily dataset on confirmed positive cases of COVID-19 in Ontario^2^. This data contains information on individual cases, including location, age, and outcome. This provided the opportunity to compute the mean death rates for each age group. For the purpose of this analysis, six cases in persons of unknown age were ignored.  

```{r mortality_calc, echo=FALSE}
deathRates = cases %>%
  filter(Outcome1 != "Not Resolved", Age_Group != "UNKNOWN") %>%
  group_by(Age_Group) %>%
  summarize(Cases = n(), Deaths = sum(Outcome1 == "Fatal"), .groups = "drop") %>%
  mutate(Death_Rate = 100*(Deaths / Cases), Age_Group = factor(Age_Group))

# Plot output
ggplot(deathRates, aes(x=Age_Group, y=Death_Rate, group=1)) +
  geom_line() +
  ggtitle("Ontario COVID-19 Mortality Rate by Age Group") +
  xlab("Age Group") +
  ylab("Death Rate (%)")

# Table output
deathRates %>%
  mutate(Death_Rate = round(Death_Rate, 4)) %>%
  rename(`Age Group` = Age_Group, `Death Rate (%)` = Death_Rate) %>%
  kable(format="simple", align="c")
```

**References:**  
^1^ "COVID-19 Case Fatality, Case Identification, and Attack Rates in Ontario.” Public Health Ontario, May 18, 2020. https://www.publichealthontario.ca/-/media/documents/ncov/epi/2020/06/covid19-epi-case-identification-age-only-template.pdf?la=en

^2^ "Confirmed Positive Cases of COVID-19 in Ontario.” Ontario Data Catalogue. Ontario Treasury Board Secretariat, March 24, 2020. https://data.ontario.ca/en/dataset/confirmed-positive-cases-of-covid-19-in-ontario