---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
```

```{r groups_pop, include=FALSE}
groups = read_csv("../data/ON_HealthUnit_Census.csv") %>%
  transmute(`Age_<20` = `Age_0-14` + `Age_15-19`,
            `Age_20s` = `Age_20-24` + `Age_25-29`,
            `Age_30s` = `Age_30-34` + `Age_35-39`,
            `Age_40s` = `Age_40-44` + `Age_45-49`,
            `Age_50s` = `Age_50-54` + `Age_55-59`,
            `Age_60s` = `Age_60-64` + `Age_65-69`,
            `Age_70s` = `Age_70-74` + `Age_75-79`,
            `Age_80+` = `Age_80-84` + `Age_Over85`) %>%
  summarize(`<20` = sum(`Age_<20`),
            `20s` = sum(`Age_20s`),
            `30s` = sum(`Age_30s`),
            `40s` = sum(`Age_40s`),
            `50s` = sum(`Age_50s`),
            `60s` = sum(`Age_60s`),
            `70s` = sum(`Age_70s`),
            `80+` = sum(`Age_80+`)) %>%
  t()

groups = tibble(`Age Group` = rownames(groups), Population = groups)
```

Given the available data for age-related mortality, we chose to divide the province's population into ten-year wide age groups. This data was collected in the 2016 Canada Census and retrieved using the CensusMapper API^1^. The populations of each age group can be seen below.  

```{r groups_out, echo=FALSE}
# Barplot
ggplot(groups, aes(x=`Age Group`, y=Population)) +
  geom_bar(stat="identity") +
  ggtitle("Population of Ontario Age Groups, 2016 Census")

# Table
kable(groups)
```

```{r infection_groups, include=FALSE}
# Get number of cases in each age group
infectByGroup = read_csv("../data/conposcovidloc.csv") %>%
  mutate(`Age Group` = ifelse(Age_Group %in% c("80s", "90s"), "80+", Age_Group)) %>%
  filter(!is.na(`Age Group`), `Age Group` != "UNKNOWN") %>%
  group_by(`Age Group`) %>%
  summarize(Cases = n())

# Compute proportion of total cases in each age group
totalCases = sum(infectByGroup$Cases)
infectByGroup = mutate(infectByGroup, Proportion = paste0(format(100*Cases / totalCases, digits=3), "%"))
```

We also considered the possibility that each age group might be at different risk levels of contracting COVID-19. We considered the proportion of confirmed COVID-19 cases in each age group, seen below.

```{r infection_plot, echo=FALSE}
infectByGroup %>%
  ggplot(aes(x=`Age Group`, y=Cases, label=Proportion)) +
  geom_bar(stat="identity") +
  geom_text(vjust=-0.5) +
  ggtitle("Ontario COVID-19 Cases by Age Group, with Proportion of Total")
```

We clearly see that some age groups contain a far greater proportion of COVID-19 cases than others. We account for this in our model by weighting new COVID-19 cases more heavily toward those groups according to the observed percentages of cases.

**References:**  
^1^ "2016 Canada Census.” https://censusmapper.ca. Statistics Canada. Accessed October 10, 2020. https://censusmapper.ca